---
title: "Income Prediction Ministry of Social Affairs Data"
author: "Mohammad Joudy"
date: "April 9, 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Main Problem

Iranian welfare center is a set of information about all of iranians which is gathered by Ministry of Labour and Social affairs in order to giving policies of the field of prosperity and other fields. This dataset is constituted from a vast spectrum of information of demographic, income, insurance, wealth and assets. 

Our dataset is a two percent sample of this huge dataset which is available in web site of ministry.

Our ultimate goal is to have a geographical and age distribution of poverty. for this purpose, we consider income of households as main feature (which would be predicted).but the problem is because of incompeleteness of this variable.

## Workflow

```{r libraries, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

rm(list = ls())

library(dplyr)
library(tidyr)
library(caret)
library(rpart)
library(DMwR)
library(ggplot2)
library(ggthemes)

set.seed(13385)
```

```{r getting data}

dataURL <- "https://www.mcls.gov.ir/icm_content/media/filepool3/2017/8/7164.rar"

Sys.setlocale(locale = "persian")

data <- read.csv("./refahdata500000-960331.csv", header = TRUE, sep = ";")
```

##Looking at Data
As we can see the dataset are comprised from 57 variables and 500,000 rows.
```{r}
str(data)
```
first of all we look at number NAs in variables. For this purpose we build a function naCountFunc.
```{r}
naCountFunc <- function(x) sum(is.na(x))
nac <- sapply(data, naCountFunc)
as.data.frame(nac)
```
So, a considerable amount of two variables named FOROSH_MAX_9504 and FEE_MAX_9504 are NAs which this is a reasonable to omit them as below.
```{r}
data <- select(data, -FOROSH_MAX_9504, -FEE_MAX_9504)
compCase <- sum(complete.cases(data))
```
now its time to split the dataset by variable of income and demography, respectively 45 and 11 columns and looking at their NAs
```{r}
daramad <- select(data, -c(2:9), -BIMEH_SALAMAT, -BIMEH_ROOSTAEIAN)
demography <- select(data, c(1:9), BIMEH_SALAMAT, BIMEH_ROOSTAEIAN)
nacDemo <- sapply(demography, naCountFunc)
as.data.frame(nacDemo)
nacD <- sapply(daramad, naCountFunc)
as.data.frame(nacD)
```
in "daramad" dataset we have different types of variables with NAs. Based on size of NAs we threat them in different ways. First, we replace NAs in varibles which their NAs are in order of 100, with their median. 

```{r}
daramad$MASKOONI_KHARID_PRICE[is.na(daramad$MASKOONI_KHARID_PRICE)] <- median(daramad$MASKOONI_KHARID_PRICE, na.rm = TRUE)
daramad$MASKOONI_FOROOSH_PRICE[is.na(daramad$MASKOONI_FOROOSH_PRICE)] <- median(daramad$MASKOONI_FOROOSH_PRICE, na.rm = TRUE)
daramad$DALIL_MODIR_BANKI[is.na(daramad$DALIL_MODIR_BANKI)] <- median(daramad$DALIL_MODIR_BANKI, na.rm = TRUE)
daramad$DALIL_ASNAF_MONTAKHAB[is.na(daramad$DALIL_ASNAF_MONTAKHAB)] <- median(daramad$DALIL_ASNAF_MONTAKHAB, na.rm = TRUE)
```
So, for variables with higher number of NAs, we apply regression trees in order to estimate them.

```{r}
modelEbrazi <- rpart(MLT_DARAMAD_EBRAZI ~ . -MLT_DARAMAD_TASHKHISI-MLT_DARAMAD_GHATEI
                     -MLT_MALIAT_TASHKHISI-MLT_MALIAT_GHATE, 
                      data = daramad[!is.na(daramad$MLT_DARAMAD_EBRAZI), ], 
                      method = "anova", na.action = na.omit)
predictEbrazi <- predict(modelEbrazi, daramad[is.na(daramad$MLT_DARAMAD_EBRAZI), ])
daramad$MLT_DARAMAD_EBRAZI[is.na(daramad$MLT_DARAMAD_EBRAZI)] <- predictEbrazi
#---------------------------------------------------------------------------------------
modelTashkhisi <- rpart(MLT_DARAMAD_TASHKHISI ~ . -MLT_DARAMAD_GHATEI
                     -MLT_MALIAT_TASHKHISI-MLT_MALIAT_GHATE, 
                     data = daramad[!is.na(daramad$MLT_DARAMAD_TASHKHISI), ], 
                     method = "anova", na.action = na.omit)
predictTashkhisi <- predict(modelTashkhisi, daramad[is.na(daramad$MLT_DARAMAD_TASHKHISI), ])
daramad$MLT_DARAMAD_TASHKHISI[is.na(daramad$MLT_DARAMAD_TASHKHISI)] <- predictTashkhisi
#---------------------------------------------------------------------------------------
modelGhatei <- rpart(MLT_DARAMAD_GHATEI ~ . -MLT_MALIAT_TASHKHISI-MLT_MALIAT_GHATE, 
                        data = daramad[!is.na(daramad$MLT_DARAMAD_GHATEI), ], 
                        method = "anova", na.action = na.omit)
predictGhatei <- predict(modelGhatei, daramad[is.na(daramad$MLT_DARAMAD_GHATEI), ])
daramad$MLT_DARAMAD_GHATEI[is.na(daramad$MLT_DARAMAD_GHATEI)] <- predictGhatei
#---------------------------------------------------------------------------------------
modelMaliatT <- rpart(MLT_MALIAT_TASHKHISI ~ . -MLT_MALIAT_GHATE, 
                     data = daramad[!is.na(daramad$MLT_MALIAT_TASHKHISI), ], 
                     method = "anova", na.action = na.omit)
predictMaliatT <- predict(modelMaliatT, daramad[is.na(daramad$MLT_MALIAT_TASHKHISI), ])
daramad$MLT_MALIAT_TASHKHISI[is.na(daramad$MLT_MALIAT_TASHKHISI)] <- predictMaliatT
#---------------------------------------------------------------------------------------
modelMaliatG <- rpart(MLT_MALIAT_GHATE ~ ., 
                      data = daramad[!is.na(daramad$MLT_MALIAT_GHATE), ], 
                      method = "anova", na.action = na.omit)
predictMaliatG <- predict(modelMaliatG, daramad[is.na(daramad$MLT_MALIAT_GHATE), ])
daramad$MLT_MALIAT_GHATE[is.na(daramad$MLT_MALIAT_GHATE)] <- predictMaliatG
```
Now, its time to do some little exploratory analysis. looking at distribution of our main variable shed light on important information.

```{r}
par(mar = rep(2,4))
boxplot(daramad$DARAMADAZHOGHOGH9504_SARPARAST)
plot(daramad$DARAMADAZHOGHOGH9504_SARPARAST)
hist(daramad$DARAMADAZHOGHOGH9504_SARPARAST, 100)
```